### 2016-03-21 Manuscript/SW_early_life_telomeres_SOM.Rmd
More analyses

#Run the model
lm1 <- lmer(RTL ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ + Season) + (1|LayYear)+(1|TerritoryID),
            data = juv,
            REML = FALSE)
lm1 <- standardize(lm1)

@@

tableS1 <- data.frame(Model = apply(t1,1,getterms,c(2:16)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS1) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS1) <- NULL
tableS1$Model[tableS1$Model == ''] <- 'Null model'
tableS1[,3:5] <- lapply(tableS1[,3:5],as.numeric)
kable(tableS1,digits=2)

@@

#Run the model
lm1 <- lmer(Loss ~ (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ + Season) + (1|LayYear),
            data = temp,
            REML = FALSE)
lm1 <- standardize(lm1)

@@

tableS2 <- data.frame(Model = apply(t1,1,getterms,c(2:8)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS2) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS2) <- NULL
tableS2$Model[tableS2$Model == ''] <- 'Null model'
tableS2[,3:5] <- lapply(tableS2[,3:5],as.numeric)
kable(tableS2,digits=2)

>>>>

#Run the model
lm1 <- lmer(RTL ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ) + (1|FieldPeriodID)+(1|TerritoryID)+(1|PlateID),
            data = juv,
            REML = FALSE)
lm1 <- standardize(lm1)

@@

tableS1 <- data.frame(Model = apply(t1,1,getterms,c(2:14)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS1) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS1) <- NULL
tableS1$Model[tableS1$Model == ''] <- 'Null model'
tableS1[,3:5] <- lapply(tableS1[,3:5],as.numeric)
kable(tableS1,digits=2)

@@

#Run the model
lm1 <- lmer(Loss ~ (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ) + (1|FieldPeriodID),
            data = temp,
            REML = FALSE)
lm1 <- standardize(lm1)

@@

tableS2 <- data.frame(Model = apply(t1,1,getterms,c(2:7)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS2) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS2) <- NULL
tableS2$Model[tableS2$Model == ''] <- 'Null model'
tableS2[,3:5] <- lapply(tableS2[,3:5],as.numeric)
kable(tableS2,digits=2)

### 2016-03-21 Manuscript/SW_early_life_telomeres_tables_and_figures.Rmd
More analyses

**Figure 1** Telomere length in relation to age in Seychelles warblers. **(A)** Mean  ($\pm$ s.e.) age for all birds. **(B)** Notched boxplot of telomere length at different age classes within the first year of life. Sample sizes for each group are provided in brackets.

**Figure 2** Factors affecting early-life telomere length in Seychelles warblers. **A** Model averaged estimates and 95% confidence intervals for explanatory terms used in a linear mixed model with telomere length (see main text for details; TQ = territory quality, TL = telomere length).

**Figure 3** Factors affecting early-life telomere shortening rates in Seychelles warblers, within seasons. Model averaged estimates and 95% confidence intervals for explanatory terms used in a linear mixed model with telomere length (cross-sectional data, **A**) and telomere loss (longitudinal data, **B**) as the response variable. Numbers in brackets are the relative importance of each term in the top model set (see main text for details; TQ = territory quality, TL = telomere length).

**Figure 4** Kaplan-Meier curves showing the relationship between survival and telomere length for juvenile Seychelles warblers. Telomere length is binned into groups here for visualisation purposes only (long and short = greater than or less than median telomere length, respectively).

@@

temp <- juv
temp$LayYear <- factor(temp$LayYear)

Fig2A <- ggplot(temp,aes(x = LayYear,y = RTL,col=Season,fill=Season))+
  ylab('RTL') + 
  xlab('Cohort') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12,angle = 90), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top')+
  geom_jitter(aes(colour = Season),position=position_jitterdodge(dodge.width = 0.8))+
  geom_boxplot(outlier.colour = NA,notch=F,alpha=0.2)+
  scale_colour_manual(values = c('grey','darkorange'))+
  scale_fill_manual(values = c('grey','darkorange'))+
  annotate('text',x = 1,y = 2.1,label = 'A')

temp <- juv
x <- ifelse(temp$Season == 'Winter',0.5,0)
temp$LayYear2 <- temp$LayYear+x
temp <- temp[!(duplicated(temp$LayYear2)),]

@@

Fig2C <- ggplot(temp,
       aes(x = Density,
           y = RTL,
           colour=Season,
           fill = Season)) +
  ylab('RTL') + 
  xlab('Adult population size') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none') +
  geom_point()+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c('grey','darkorange','darkgreen'))+
  scale_fill_manual(values = c('grey','darkorange','darkgreen'))+
  annotate('text',x = 240,y = 2.1,label = 'C')

pdf('Figures/Figure 2.pdf',
    width = 8,
    height = 8,
    paper = 'a4')
multiplot(Fig2A,Fig2B,Fig2C,layout=matrix(c(1,1,2,3),nrow = 2,byrow=T))
dev.off()

@@

#####################
#2A
#####################

#Run the model
lm2 <- lmer(RTL ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ + Season)  + (1|LayYear) + (1|TerritoryID),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)
#Model averaging

@@

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex (males)'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'c.Season'] <- 'Season (winter)'
nams[nams == 'c.Season:z.Agemonths'] <- 'Age x season'
nams[nams == 'z.Agemonths:z.NonHelper'] <- 'Age x non helpers'
nams[nams == 'c.Sex:z.Agemonths'] <- 'Age x sex'
nams[nams == 'z.Agemonths:z.TQ'] <- 'Age x TQ'
nams[nams == 'c.Helper:z.Agemonths'] <- 'Age x helpers'
nams[nams == 'z.Agemonths:z.BodyMass'] <- 'Age x mass'
nams[nams == 'z.Agemonths:z.Tarsus'] <- 'Age x tarsus'

@@

Fig3B <- ggplot(temp,
       aes(x = `Age (months)`,
           y = RTL,
           fill=Season)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_jitter(aes(colour = Season),position=position_jitterdodge(dodge.width = 0.8))+
  geom_boxplot(outlier.colour = NA,notch=F,alpha=0.2)+
  scale_colour_manual(values = c('grey','darkorange'))+
  scale_fill_manual(values = c('grey','darkorange'))+
  annotate('text',x = 0.55,y = 2.1,label = 'B')

Fig3C <- ggplot(temp,
       aes(x = Tarsus,
           y = RTL,
           colour=`Age (months)`,
           fill = `Age (months)`,
           shape = `Age (months)`)) +
  ylab('RTL') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_point()+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c(grey(0.2),'darkorange','darkgreen'))+
  scale_fill_manual(values = c('darkgrey','darkorange','darkgreen'))+
  annotate('text',x = 17,y = 2.1,label = 'C')

Fig3D <- ggplot(temp,
       aes(x = `Age (months)`,
           y = RTL,
           fill=Sex)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_jitter(aes(colour = Sex),position=position_jitterdodge(dodge.width = 0.8))+
  geom_boxplot(outlier.colour = NA,notch=F,alpha=0.2)+
  scale_colour_manual(values = c('grey','darkorange'))+
  scale_fill_manual(values = c('grey','darkorange'))+
  annotate('text',x = 0.55,y = 2.1,label = 'D')

pdf('Figures/Figure 3.pdf',
    width = 9,
    height = 9,
    paper = 'a4r')
multiplot(Fig3A,Fig3B,Fig3C,Fig3D,layout = matrix(c(1,1,1,2,3,4),nrow = 2,byrow = T))
dev.off()

@@

Fig4B <- ggplot(temp2,
       aes(x = TimeDiff,
           y = Loss,
           colour=`Age (months)`,
           fill = `Age (months)`)) +
  ylab(expression(Delta*'RTL')) + 
  xlab('Time (years)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_point(alpha=0)+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c(grey(0.2),'darkorange','darkgreen'))+
  scale_fill_manual(values = c('darkgrey','darkorange','darkgreen'))+
    geom_abline(slope = 0,intercept = 0,lty = 2)+
  coord_cartesian(ylim = c(-0.5,0.25))+
  annotate('text',x = 0,y = 0.5,label = 'B')

@@

#####################
#4A
#####################
temp <- subset(Loss,Agemonths == 1)
#Run the model
lm2 <- lmer(Loss ~ (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ) + (1|LayYear),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)

@@

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'c.Season'] <- 'Season'

@@

Fig4 <- forestplot(table2)

pdf('Figures/Figure 4.pdf',
    width = 5,
    height = 4,
    paper = 'a4')
Fig5
dev.off()

>>>>

**Figure 1** RTL in relation to **(A)** age and **(B)** $\Delta$Age in Seychelles warblers. Lines and shaded areas are fitted values and 95% confidence limits from linear regressions.

**Figure 2** Cohort effects on early-life RTL in Seychelles warblers. **A** temporal variation RTL in major and minor cohorts, **B** tempoeral variation in adult population size, and **C** early-life RTL in relation to cohort-level variation in adult population size. Lines and shaded areas in **C** are fitted values and 95% confidence limits from general linear models.

**Figure 3** Factors affecting early-life RTL in Seychelles warblers. **A** Model averaged estimates and 95% confidence intervals for explanatory terms used in a linear mixed model with RTL as the response variable (see main text for details; TQ = territory quality). **B-D** early life RTL in relation to season (**B**),tarsus length (**C**) and sex (**D**). Lines and shaded areas in **D** are fitted values and 95% confidence limits from general linear models.

**Figure** Longitudinal telomere dynamics in juvenile Seychelles warblers. **A** Juvenile telomere length in relation to adult telomere length within individuals. **B** $\Delta$RTL in relation to the time between sampling events. Lines and shaded areas represent fitted values and 95% confidence limits from general linear models. Raw data have been ommited from **B** for the sake of clarity, but are plotted in Figure S1.

**Figure 5** Factors affecting $\Delta$RTL in juvenile Seychelles warblers. Model averaged estimates and 95% confidence intervals for explanatory terms used in a linear mixed model with $\Delta$RTL as the response variable (see main text for details; TQ = territory quality).

@@

temp <- juv
x <- ifelse(temp$Season == 'Minor',0.5,0)
temp$LayYear2 <- temp$LayYear+x
temp$LayYear <- factor(temp$LayYear)

Fig2A <- ggplot(temp,aes(x = LayYear,y = RTL,fill=Season))+
  ylab('RTL') + 
  xlab('Cohort') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12,angle = 90), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top')+
  geom_jitter(aes(colour = Season),position=position_jitterdodge(dodge.width = 0.8))+
  geom_boxplot(outlier.colour = NA,notch=F,alpha=0.2)+
  scale_colour_manual(values = c('grey','darkorange'))+
  scale_fill_manual(values = c('grey','darkorange'))+
  annotate('text',x = 1,y = 2.1,label = 'A')

temp <- temp[!(duplicated(temp$LayYear2)),]

@@

temp <- juv
temp[,'Age (months)'] <- factor(temp$Agemonths)

Fig2C <- ggplot(temp,
       aes(x = Density,
           y = RTL,
           colour=Season,
           fill = Season)) +
  ylab('RTL') + 
  xlab('Adult population size') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none') +
  geom_point()+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c('grey','darkorange','darkgreen'))+
  scale_fill_manual(values = c('grey','darkorange','darkgreen'))+
  annotate('text',x = 250,y = 2.1,label = 'C')

pdf('Figures/Figure 2.pdf',
    width = 8,
    height = 8,
    paper = 'a4r')
multiplot(Fig2A,Fig2B,Fig2C,layout=matrix(c(1,1,2,3),nrow = 2,byrow=T))
dev.off()

@@

#Run the model
lm2 <- lmer(RTL ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ)  + (1|FieldPeriodID) + (1|TerritoryID)+(1|PlateID),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)
#Model averaging

@@

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex (males)'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'c.Season'] <- 'Season (minor)'
nams[nams == 'c.Season:z.Agemonths'] <- 'Age x season'
nams[nams == 'z.Agemonths:z.NonHelper'] <- 'Age x non helpers'
nams[nams == 'c.Sex:z.Agemonths'] <- 'Age x sex'
nams[nams == 'z.Agemonths:z.TQ'] <- 'Age x TQ'
nams[nams == 'c.Helper:z.Agemonths'] <- 'Age x helpers'
nams[nams == 'z.Agemonths:z.BodyMass'] <- 'Age x mass'
nams[nams == 'z.Agemonths:z.Tarsus'] <- 'Age x tarsus'

@@

Fig3B <- ggplot(temp,
       aes(x = Tarsus,
           y = RTL,
           colour=`Age (months)`,
           fill = `Age (months)`,
           shape = `Age (months)`)) +
  ylab('RTL') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_point()+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c('grey','darkorange','darkgreen'))+
  scale_fill_manual(values = c('grey','darkorange','darkgreen'))+
  annotate('text',x = 17,y = 2.1,label = 'B')

Fig3C <- ggplot(temp,
       aes(x = `Age (months)`,
           y = RTL,
           fill=Sex)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_jitter(aes(colour = Sex),position=position_jitterdodge(dodge.width = 0.8))+
  geom_boxplot(outlier.colour = NA,notch=F,alpha=0.2)+
  scale_colour_manual(values = c('grey','darkorange'))+
  scale_fill_manual(values = c('grey','darkorange'))+
  annotate('text',x = 0.55,y = 2.1,label = 'C')

pdf('Figures/Figure 3.pdf',
    width = 8,
    height = 8,
    paper = 'a4r')
multiplot(Fig3A,Fig3B,Fig3C,layout = matrix(c(1,1,2,3),nrow = 2,byrow = T))
dev.off()

@@

Fig4B <- ggplot(temp2,
       aes(x = TimeDiff,
           y = Loss,
           colour=`Age (months)`,
           fill = `Age (months)`)) +
  ylab(expression(Delta*'RTL')) + 
  xlab('Time (years)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_point(alpha=0)+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c('grey','darkorange','darkgreen'))+
  scale_fill_manual(values = c('grey','darkorange','darkgreen'))+
    geom_abline(slope = 0,intercept = 0,lty = 2)+
  coord_cartesian(ylim = c(-0.5,0.3))+
  annotate('text',x = 0,y = 0.25,label = 'B')

@@

temp <- subset(Loss,Agemonths == 1)
#Run the model
lm2 <- lmer(Loss ~ (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ) + (1|FieldPeriodID),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)

@@

#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex (males)'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'

@@

Fig5A <- forestplot(table2)+
  annotate('text',x = 0.75,y = 0.45,label = 'A')

Fig5B <- ggplot(temp,
       aes(x = Tarsus,
           y = Loss,
           col=Sex)) +
  ylab(expression(Delta*'RTL')) + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'top') +
  geom_point()+
  stat_smooth(method = 'lm')+
  scale_colour_manual(values = c('grey','darkorange','darkgreen'))+
  scale_fill_manual(values = c('grey','darkorange','darkgreen'))+
  annotate('text',x = 19,y = 1.1,label = 'B')

pdf('Figures/Figure 5.pdf',
    width = 4,
    height = 8,
    paper = 'a4')
multiplot(Fig5A,Fig5B)
dev.off()
